---
title: "Project R"
author: "Julien Legrand"
date: "2025-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project NYC Airbnb - Analyzing the Rental Market

Your data analyst team has been hired to study the Airbnb market in New York City. Using a public dataset, you must analyze price trends, the geographic distribution of listings, and the factors that influence rental prices. The project will culminate in the creation of a regression model to predict the price of a rental and an interactive map to visualize your findings.

```{r, include=FALSE}
dataset = read.csv("train.csv")
library(lubridate)
library(dplyr)
```

Handling Temporal Data: The last_review column is in text format. Convert it to a proper date format using the lubridate package.

```{r}
head(dataset)
```

```{r}
dataset$last_review = ymd(dataset$last_review)
dataset
```

 Treating Informative NA s: The last_review and reviews_per_month columns have NA s for listings with no reviews. Replace the .NA s in 
reviews_per_month with a proper number.

```{r}
dataset$reviews_per_month[is.na(dataset$reviews_per_month)] =0 
head(dataset)
```

```{r}
clean_dataset = dataset [dataset$price != 0,]
clean_dataset
```

```{r}

```

