---
title: "Project R"
author: "Legrand, Hu, Stintzi"
date: "2025-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project NYC Airbnb - Analyzing the Rental Market

Your data analyst team has been hired to study the Airbnb market in New York City. Using a public dataset, you must analyze price trends, the geographic distribution of listings, and the factors that influence rental prices. The project will culminate in the creation of a regression model to predict the price of a rental and an interactive map to visualize your findings.

```{r, include=FALSE}
train = read.csv("train.csv")
library(lubridate)
library(dplyr)
```

Handling Temporal Data: The last_review column is in text format. Convert it to a proper date format using the lubridate package.

```{r}
head(train)
```

```{r}
train$last_review = ymd(train$last_review)
train
```

Treating Informative NA s: The last_review and reviews_per_month columns have NA s for listings with no reviews. Replace the .NA s in reviews_per_month with a proper number.

```{r}
train$reviews_per_month[is.na(train$reviews_per_month)] = 0
train
```

```{r}
CleanTrainPrice = train %>%
  filter(price != 0)
CleanTrainPrice
```

```{r}
Clean_train = CleanTrainPrice %>%
  select(-c(id, host_id, host_name))
Clean_train
```

## Step 2:

```{r}
Clean_train = Clean_train %>%
  mutate(price_log = log1p(price))
```

Verify the difference between price and price_log

```{r}
summary(Clean_train$price)
summary(Clean_train$price_log)
```

```{r}
# Calcul de la date la plus récente dans le dataset
max_date = max(Clean_train$last_review, na.rm = TRUE)

# Création de la nouvelle colonne
Clean_train = Clean_train %>%
  mutate(days_since_last_review = as.numeric(difftime(max_date, last_review, units = "days")))

# Vérification
Clean_train
```

```{r}
# Aperçu de la répartition des quartiers
table(Clean_train$neighbourhood_group)

# Et du nombre de quartiers détaillés
length(unique(Clean_train$neighbourhood))
```

```{r}
Clean_train$room_type = as.factor(Clean_train$room_type)

str(Clean_train$room_type)
levels(Clean_train$room_type)
```

## Step 3

```{r, include = FALSE}
library(leaflet)
library(ggplot2)
library(forcats)
```

### Creation of the interactive map
```{r}
#We instanciate color for the neighbourhood
interactive_map_color <- colorFactor(c("red", "blue", "yellow", "green", "purple"), domain = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"))

#We create the map
interactive_map = leaflet()
#We had the layer of the map
interactive_map <- addTiles(interactive_map)
interactive_map <- addProviderTiles(
  map = interactive_map,
  provider = providers$CartoDB.Positron
)

#We add the markers for every AirBNB
interactive_map <- addCircleMarkers(
  map = interactive_map,
  lng = Clean_train$longitude,
  lat = Clean_train$latitude,
  popup = Clean_train$name,
  label = Clean_train$name,
  color = interactive_map_color(Clean_train$neighbourhood_group),
  fillColor = interactive_map_color(Clean_train$neighbourhood_group), 
  fillOpacity = 0.8,
  radius = Clean_train$price_log / 100
)

#We instanciate the legend
interactive_map <- addLegend(
  map = interactive_map,
  position = "topright",
  pal = interactive_map_color,
  values = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"), 
  title = "Neighbourhood",
  opacity = 1
  
)


```

```{r}
interactive_map
```
Because there's more than 44.000 AirBnB in New York City, the map lag in the area.


### Price distribution by borought
```{r}
#We calculate the mean price for every neighbourhood
Mean_neighbourhood = Clean_train %>%
  group_by(neighbourhood) %>%
  summarise(Mean_price = mean(price_log, na.rm = TRUE))

#We add the neighbourhood_group in the new dataset
neighbourhood_group_reference <- Clean_train %>%
  select(neighbourhood, neighbourhood_group) %>%
  distinct()
Mean_complete_neighbourhood <- Mean_neighbourhood %>%
  left_join(neighbourhood_group_reference, by = "neighbourhood")

#We plot the result
plot_distribution_box = ggplot(Mean_complete_neighbourhood) + 
  aes(x = neighbourhood_group, y = Mean_price, colour = neighbourhood_group) + 
  geom_jitter(size = 1.5, alpha = 1, aes(colour = neighbourhood_group), fill = NA) + 
  geom_boxplot(colour = "gray60", size=1, alpha = 0.5) + 
  labs(title = "Mean prices of the differents neighbourhood", x="neighbourhood", y = "Mean prices") + 
  theme_minimal() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) 

plot_distribution_box
```
On this graphic, the points are the mean price of each neighbourhood. The boxplots are the price mean of every neighbourhood_group, based on the price mean of the neighbourhood

### Distribution of listing type
```{r}
#We take the sum of listing for each Room type
Number_of_listing = Clean_train %>%
  group_by(room_type) %>%
  summarise(nmb_listing = sum(calculated_host_listings_count))

#We plot the results
ggplot(Number_of_listing, aes(x=room_type, y=nmb_listing, color=room_type, fill= room_type))+geom_col()+labs(title= "Number of listing per room type", x= "Room type", y= "Number of listing") + theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10))

```
As we can see, most of the AirBnB booked are the entire Home / apartment. This is quite logical because it's what AirBnB push on

### Relationship between availability and Reviews
```{r}
#We plot the relationship
ggplot(Clean_train, aes(x=availability_365, y = number_of_reviews, color=neighbourhood_group)) + geom_point() + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10))
```
As we can see, there's too much element to be clear. We will now separate by neighbourhood
```{r}
#We plot the results

#For Bronx
ggplot(Clean_train[Clean_train$neighbourhood_group == "Bronx",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="red", size= 0.75) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Bronx") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))

#For Brooklyn
ggplot(Clean_train[Clean_train$neighbourhood_group == "Brooklyn",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="yellow3", size= 0.75) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Brooklyn") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))

#For Manhattan
ggplot(Clean_train[Clean_train$neighbourhood_group == "Manhattan",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="green3", size= 0.75, alpha = 0.1) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Manhattan") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))

#For Queens
ggplot(Clean_train[Clean_train$neighbourhood_group == "Queens",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="blue", size= 0.75) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Queens") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))

#For Staten Island
ggplot(Clean_train[Clean_train$neighbourhood_group == "Staten Island",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="purple", size= 0.75) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Staten Island") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))
```
As we can see, it's still not enough to have a good view on all the data we want, especially in Brooklyn, Manhattan and Queens. So we will cut the graphs in 2, in order to make it easier to read.

```{r }
#For Bronx
ggplot(Clean_train[Clean_train$neighbourhood_group == "Bronx",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="red", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Bronx for availability between 0 to 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190), ylim = c(0, 350))


ggplot(Clean_train[Clean_train$neighbourhood_group == "Bronx",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="red", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Bronx for availability between 190 and 380 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190,380), ylim = c(0, 350))
```
```{r}
#For Brooklyn
ggplot(Clean_train[Clean_train$neighbourhood_group == "Brooklyn",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="yellow3", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Brooklyn for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190),ylim = c(0, 450))

ggplot(Clean_train[Clean_train$neighbourhood_group == "Brooklyn",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="yellow3", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Brooklyn for availability between 190 and 380 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190,380), ylim = c(0, 450))
```

```{r}
#For Manhattan
ggplot(Clean_train[Clean_train$neighbourhood_group == "Manhattan",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="green3", size= 0.75, alpha = 1) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Manhattan for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190),ylim = c(0, 300))

ggplot(Clean_train[Clean_train$neighbourhood_group == "Manhattan",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="green3", size= 0.75, alpha = 1) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Manhattan for availability between 190 and 380 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190,380), ylim = c(0, 300))
```

```{r}
#For Queens
ggplot(Clean_train[Clean_train$neighbourhood_group == "Queens",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="blue", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Queens for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190), ylim = c(0, 450))

ggplot(Clean_train[Clean_train$neighbourhood_group == "Queens",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="blue", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Queens for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190,380), ylim = c(0, 450))
```

```{r}
#For Staten Island
ggplot(Clean_train[Clean_train$neighbourhood_group == "Staten Island",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="purple", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Staten Island for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190), ylim = c(0, 450))

ggplot(Clean_train[Clean_train$neighbourhood_group == "Staten Island",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="purple", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Staten Island for availability between 190 and 380 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190, 380), ylim = c(0, 450))
```

Now, the graphs looks much easier to understand.


### Comparaison entre price et minimum night

```{r}
#We plot the results
ggplot(Clean_train, aes(x = minimum_nights, y = price, color = "red")) + geom_point() + theme_minimal() + labs(x = "Number of nights required", y = "Price of the AirBnB", title = "The relationship between the price and the number of nights") + theme(legend.position = "none",plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) 

```
As we can see, some AirBnB asks for too much nights, which makes it not interesting (and not possible). If we were staff members form AirBnB, we would have looked at them in order to be sure everything is alright with theses announcement.
We will now look much closer on those who requires less than a year of minimum nights, in order to have a better view of them.

```{r}
#We plot the results
ggplot(Clean_train, aes(x = minimum_nights, y = price, color = "red")) + geom_point(size = 0.5) + theme_minimal() + labs(x = "Number of nights required", y = "Price of the AirBnB", title = "The relationship between the price and the number of nights") + theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0, 365))
```

And now, the relationships for the flats that requires less than 100 days in it :

```{r}
#We plot the result
ggplot(Clean_train, aes(x = minimum_nights, y = price, color = "red")) + geom_point(size = 0.5) + theme_minimal() + labs(x = "Number of nights required", y = "Price of the AirBnB", title = "The relationship between the price and the number of nights") + theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0, 100), ylim = c(0,6000))
```

### Correlation between the number of review and the price of AirBnB

```{r}
#We plot the results
ggplot(Clean_train, aes(x = number_of_reviews, y = price, color = "red")) + geom_point() + theme_minimal() + labs(x = "Number of reviews", y = "Price of the AirBnB", title = "The relationship between the price and the number of reviews") + theme(legend.position = "none",plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) 
```
As we can see, the less the price is, the more reviews there is. Cheap AirBnB are more attractive.
