---
title: "Project R"
author: "Legrand, Hu, Stintzi"
date: "2025-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project NYC Airbnb - Analyzing the Rental Market

Your data analyst team has been hired to study the Airbnb market in New York City. Using a public dataset, you must analyze price trends, the geographic distribution of listings, and the factors that influence rental prices. The project will culminate in the creation of a regression model to predict the price of a rental and an interactive map to visualize your findings.

```{r, include=FALSE}
train = read.csv("train.csv")
library(lubridate)
library(dplyr)
```

Handling Temporal Data: The last_review column is in text format. Convert it to a proper date format using the lubridate package.

```{r}
head(train)
```

```{r}
train$last_review = ymd(train$last_review)
train
```

 Treating Informative NA s: The last_review and reviews_per_month columns have NA s for listings with no reviews. Replace the .NA s in 
reviews_per_month with a proper number.

```{r}
train$reviews_per_month[is.na(train$reviews_per_month)] = 0
train
```

```{r}
CleanTrainPrice = train %>%
  filter(price != 0)
CleanTrainPrice
```

```{r}
Clean_train = CleanTrainPrice %>%
  select(-c(id, host_id, host_name))
Clean_train
```

## Step 2:

```{r}
Clean_train = Clean_train %>%
  mutate(price_log = log1p(price))
```


Verify the difference between price and price_log

```{r}
summary(Clean_train$price)
summary(Clean_train$price_log)
```

```{r}
# Calcul de la date la plus récente dans le dataset
max_date = max(Clean_train$last_review, na.rm = TRUE)

# Création de la nouvelle colonne
Clean_train = Clean_train %>%
  mutate(days_since_last_review = as.numeric(difftime(max_date, last_review, units = "days")))

# Vérification
Clean_train
```

```{r}
# Aperçu de la répartition des quartiers
table(Clean_train$neighbourhood_group)

# Et du nombre de quartiers détaillés
length(unique(Clean_train$neighbourhood))
```

```{r}
Clean_train$room_type = as.factor(Clean_train$room_type)

str(Clean_train$room_type)
levels(Clean_train$room_type)
```

## Step 3

```{r}
library(leaflet)
library(ggplot2)
library(forcats)
```

```{r}
interactive_map = leaflet()
interactive_map <- addTiles(interactive_map)

interactive_map_color <- colorFactor(c("red", "blue", "yellow", "green", "purple"), domain = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"))

interactive_map <- addProviderTiles(
  map = interactive_map,
  provider = providers$CartoDB.Positron
)
interactive_map <- addCircleMarkers(
  map = interactive_map,
  lng = Clean_train$longitude,
  lat = Clean_train$latitude,
  popup = Clean_train$name,
  label = Clean_train$name,
  color = interactive_map_color(Clean_train$neighbourhood_group),
  fillColor = interactive_map_color(Clean_train$neighbourhood_group), 
  fillOpacity = 0.8,
  radius = Clean_train$price_log / 100
)
interactive_map <- addLegend(
  map = interactive_map,
  position = "topright",
  pal = interactive_map_color,
  values = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"), 
  title = "Quartier",
  opacity = 1
  
)


```

```{r}
interactive_map
```


```{r}
Mean_neighbourhood = Clean_train %>%
  group_by(neighbourhood) %>%
  summarise(Mean_price = mean(price_log, na.rm = TRUE))

#Add the neighbourhood group in the dataset
neighbourhood_group_reference <- Clean_train %>%
  select(neighbourhood, neighbourhood_group) %>%
  distinct()
Mean_complete_neighbourhood <- Mean_neighbourhood %>%
  left_join(neighbourhood_group_reference, by = "neighbourhood")

plot_distribution_box = ggplot(Mean_complete_neighbourhood) + 
  aes(x = neighbourhood_group, y = Mean_price, colour = neighbourhood_group) + 
  geom_jitter(size = 1.5, alpha = 1, aes(colour = neighbourhood_group), fill = NA) + 
  geom_boxplot(colour = "gray60", size=1, alpha = 0.5) + 
  labs(title = "Mean prices of the differents neighbourhood", x="neighbourhood", y = "Mean prices") + 
  theme_minimal() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) 
Mean_complete_neighbourhood
plot_distribution_box
```

```{r}
Number_of_listing = Clean_train %>%
  group_by(room_type) %>%
  summarise(nmb_listing = sum(calculated_host_listings_count))

ggplot(Number_of_listing, aes(x=room_type, y=nmb_listing, color=room_type, fill= room_type))+geom_col()+labs(title= "Number of listing per room type", x= "Room type", y= "Number of listing")
```


