---
title: "Project R"
author: "Legrand, Hu, Stintzi"
date: "2025-10-22"
output:
  html_document:
    css: ProjetR.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Project NYC Airbnb - Analyzing the Rental Market

### Projet Goal :

Your data analyst team has been hired to study the Airbnb market in New York City. Using a public dataset, you must analyze price trends, the geographic distribution of listings, and the factors that influence rental prices. The project will culminate in the creation of a regression model to predict the price of a rental and an interactive map to visualize your findings.

## Step 1 : Data Cleaning and Preparation

-   Handling Temporal Data: The column last_review has been converted to date format in order to be able to perform temporal analyses.

-   Treating Informative NAs: The missing values in reviews_per_month have been replaced with 0, as they correspond to homes without reviews.

-   Cleaning the Price Column: Ads with a price equal to 0 have been removed, because they are not realistic.

-   Removing Irrelevant Columns: The id, host_id and host_name columns have been removed because they are not useful for general analysis.

```{r, include=FALSE}
train = read.csv("train.csv")
library(lubridate)
library(dplyr)
library(knitr)
```

Handling Temporal Data: The last_review column is in text format. Convert it to a proper date format using the lubridate package.

```{r, include=FALSE}
train$last_review = ymd(train$last_review)
train
```

Treating Informative NA s: The last_review and reviews_per_month columns have NA s for listings with no reviews. Replace the .NA s in reviews_per_month with a proper number.

```{r, include=FALSE}
train$reviews_per_month[is.na(train$reviews_per_month)] = 0
train
```

```{r, include=FALSE}
CleanTrainPrice = train %>%
  filter(price != 0)
CleanTrainPrice
```

```{r, include=FALSE}
Clean_train = CleanTrainPrice %>%
  select(-c(id, host_id, host_name))
Clean_train
```

```{r}
kable(head(Clean_train), caption = "Aperçu du jeu de données nettoyé")
```


## Step 2 : Feature Transformation and Engineering

### Logarithmic Price Transformation

-   The variable price was transformed into price_log using a logarithmic transformation

```{r}
Clean_train = Clean_train %>%
  mutate(price_log = log1p(price))
```

Verify the difference between price and price_log

```{r}
summary(Clean_train$price)
summary(Clean_train$price_log)
```

### Creating a Recency Feature

-   A new variable days_since_last_review has been created to know how many days have passed since the last review

```{r, include=FALSE}
# Calcul de la date la plus récente dans le dataset
max_date = max(Clean_train$last_review, na.rm = TRUE)

# Création de la nouvelle colonne
Clean_train = Clean_train %>%
  mutate(days_since_last_review = as.numeric(difftime(max_date, last_review, units = "days")))

# Vérification
Clean_train
```

### Simplifying Neighborhoods

-   The neighborhood analysis was checked to identify the most suitable column for visualizations.

```{r}
# Aperçu de la répartition des quartiers
table(Clean_train$neighbourhood_group)

# Et du nombre de quartiers détaillés
length(unique(Clean_train$neighbourhood))
```

### Encoding Room Types

-   The room_type variable has been converted into a factor so that it is recognized as a categorical variable by R.

```{r}
Clean_train$room_type = as.factor(Clean_train$room_type)

str(Clean_train$room_type)
levels(Clean_train$room_type)
```

## Step 3 : Visualization and Exploration

```{r, include = FALSE}
library(leaflet)
library(ggplot2)
library(forcats)
```

### Creation of the interactive map

-   

```{r}
#We instanciate color for the neighbourhood
interactive_map_color <- colorFactor(c("red", "blue", "yellow", "green", "purple"), domain = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"))

#We create the map
interactive_map = leaflet()
#We had the layer of the map
interactive_map <- addTiles(interactive_map)
interactive_map <- addProviderTiles(
  map = interactive_map,
  provider = providers$CartoDB.Positron
)

#We add the markers for every AirBNB
interactive_map <- addCircleMarkers(
  map = interactive_map,
  lng = Clean_train$longitude,
  lat = Clean_train$latitude,
  popup = Clean_train$name,
  label = Clean_train$name,
  color = interactive_map_color(Clean_train$neighbourhood_group),
  fillColor = interactive_map_color(Clean_train$neighbourhood_group), 
  fillOpacity = 0.8,
  radius = Clean_train$price_log / 100
)

#We instanciate the legend
interactive_map <- addLegend(
  map = interactive_map,
  position = "topright",
  pal = interactive_map_color,
  values = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"), 
  title = "Neighbourhood",
  opacity = 1
  
)


```

```{r}
interactive_map
```

Because there's more than 44.000 AirBnB in New York City, the map lag in the area.

### Price distribution by borought

-   The distribution of prices (on the logarithmic scale) was compared between the different neighborhoods to visualize cost differences.

```{r}
#We calculate the mean price for every neighbourhood
Mean_neighbourhood = Clean_train %>%
  group_by(neighbourhood) %>%
  summarise(Mean_price = mean(price_log, na.rm = TRUE))

#We add the neighbourhood_group in the new dataset
neighbourhood_group_reference <- Clean_train %>%
  select(neighbourhood, neighbourhood_group) %>%
  distinct()
Mean_complete_neighbourhood <- Mean_neighbourhood %>%
  left_join(neighbourhood_group_reference, by = "neighbourhood")

#We plot the result
plot_distribution_box = ggplot(Mean_complete_neighbourhood) + 
  aes(x = neighbourhood_group, y = Mean_price, colour = neighbourhood_group) + 
  geom_jitter(size = 1.5, alpha = 1, aes(colour = neighbourhood_group), fill = NA) + 
  geom_boxplot(colour = "gray60", size=1, alpha = 0.5) + 
  labs(title = "Mean prices of the differents neighbourhood", x="neighbourhood", y = "Mean prices") + 
  theme_minimal() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) 

plot_distribution_box
```

On this graphic, the points are the mean price of each neighbourhood. The boxplots are the price mean of every neighbourhood_group, based on the price mean of the neighbourhood

### Distribution of listing type

-   Distribution of Listing Types: A graph shows the number of dwellings by room type (room_type), revealing that the majority of rentals are apartments or entire houses.

```{r}
#We take the sum of listing for each Room type
Number_of_listing = Clean_train %>%
  group_by(room_type) %>%
  summarise(nmb_listing = sum(calculated_host_listings_count))

#We plot the results
ggplot(Number_of_listing, aes(x=room_type, y=nmb_listing, color=room_type, fill= room_type))+geom_col()+labs(title= "Number of listing per room type", x= "Room type", y= "Number of listing") + theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10))

```

As we can see, most of the AirBnB booked are the entire Home / apartment. This is quite logical because it's what AirBnB push on

### Relationship between availability and Reviews

-   Graphs explore the relationship between availability (availability_365) and number of reviews (number_of_reviews) for each borough, facilitating the identification of trends.

```{r}
#We plot the relationship
ggplot(Clean_train, aes(x=availability_365, y = number_of_reviews, color=neighbourhood_group)) + geom_point() + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10))
```

As we can see, there's too much element to be clear. We will now separate by neighbourhood

```{r}
#We plot the results

#For Bronx
ggplot(Clean_train[Clean_train$neighbourhood_group == "Bronx",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="red", size= 0.75) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Bronx") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))

#For Brooklyn
ggplot(Clean_train[Clean_train$neighbourhood_group == "Brooklyn",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="yellow3", size= 0.75) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Brooklyn") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))

#For Manhattan
ggplot(Clean_train[Clean_train$neighbourhood_group == "Manhattan",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="green3", size= 0.75, alpha = 0.1) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Manhattan") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))

#For Queens
ggplot(Clean_train[Clean_train$neighbourhood_group == "Queens",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="blue", size= 0.75) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Queens") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))

#For Staten Island
ggplot(Clean_train[Clean_train$neighbourhood_group == "Staten Island",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="purple", size= 0.75) + theme_minimal() + labs(x="Availability on 365 days", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Staten Island") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(ylim = c(0, 450))
```

As we can see, it's still not enough to have a good view on all the data we want, especially in Brooklyn, Manhattan and Queens. So we will cut the graphs in 2, in order to make it easier to read.

```{r }
#For Bronx
ggplot(Clean_train[Clean_train$neighbourhood_group == "Bronx",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="red", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Bronx for availability between 0 to 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190), ylim = c(0, 350))


ggplot(Clean_train[Clean_train$neighbourhood_group == "Bronx",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="red", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Bronx for availability between 190 and 380 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190,380), ylim = c(0, 350))
```

```{r}
#For Brooklyn
ggplot(Clean_train[Clean_train$neighbourhood_group == "Brooklyn",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="yellow3", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Brooklyn for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190),ylim = c(0, 450))

ggplot(Clean_train[Clean_train$neighbourhood_group == "Brooklyn",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="yellow3", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Brooklyn for availability between 190 and 380 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190,380), ylim = c(0, 450))
```

```{r}
#For Manhattan
ggplot(Clean_train[Clean_train$neighbourhood_group == "Manhattan",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="green3", size= 0.75, alpha = 1) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Manhattan for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190),ylim = c(0, 300))

ggplot(Clean_train[Clean_train$neighbourhood_group == "Manhattan",], aes(x=availability_365, y = number_of_reviews)) + geom_point( color="green3", size= 0.75, alpha = 1) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Manhattan for availability between 190 and 380 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190,380), ylim = c(0, 300))
```

```{r}
#For Queens
ggplot(Clean_train[Clean_train$neighbourhood_group == "Queens",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="blue", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Queens for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190), ylim = c(0, 450))

ggplot(Clean_train[Clean_train$neighbourhood_group == "Queens",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="blue", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Queens for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190,380), ylim = c(0, 450))
```

```{r}
#For Staten Island
ggplot(Clean_train[Clean_train$neighbourhood_group == "Staten Island",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="purple", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Staten Island for availability between 0 and 190 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0,190), ylim = c(0, 450))

ggplot(Clean_train[Clean_train$neighbourhood_group == "Staten Island",], aes(x=availability_365, y = number_of_reviews)) + geom_point(color="purple", size= 0.75) + theme_minimal() + labs(x="Availability", y = "Number of Reviews", title = "Number of reviews correlated to the availability in Staten Island for availability between 190 and 380 days") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(190, 380), ylim = c(0, 450))
```

Now, the graphs looks much easier to understand.

### Comparaison entre price et minimum night

-   Additional charts analyze the relationship between price and minimum nights, as well as price and number of reviews, to better understand the Airbnb market.

```{r}
#We plot the results
ggplot(Clean_train, aes(x = minimum_nights, y = price, color = "red")) + geom_point() + theme_minimal() + labs(x = "Number of nights required", y = "Price of the AirBnB", title = "The relationship between the price and the number of nights") + theme(legend.position = "none",plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) 

```

As we can see, some AirBnB asks for too much nights, which makes it not interesting (and not possible). If we were staff members form AirBnB, we would have looked at them in order to be sure everything is alright with theses announcement. We will now look much closer on those who requires less than a year of minimum nights, in order to have a better view of them.

```{r}
#We plot the results
ggplot(Clean_train, aes(x = minimum_nights, y = price, color = "red")) + geom_point(size = 0.5) + theme_minimal() + labs(x = "Number of nights required", y = "Price of the AirBnB", title = "The relationship between the price and the number of nights") + theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0, 365))
```

And now, the relationships for the flats that requires less than 100 days in it :

```{r}
#We plot the result
ggplot(Clean_train, aes(x = minimum_nights, y = price, color = "red")) + geom_point(size = 0.5) + theme_minimal() + labs(x = "Number of nights required", y = "Price of the AirBnB", title = "The relationship between the price and the number of nights") + theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) + coord_cartesian(xlim = c(0, 100), ylim = c(0,6000))
```

### Correlation between the number of review and the price of AirBnB

```{r}
#We plot the results
ggplot(Clean_train, aes(x = number_of_reviews, y = price, color = "red")) + geom_point() + theme_minimal() + labs(x = "Number of reviews", y = "Price of the AirBnB", title = "The relationship between the price and the number of reviews") + theme(legend.position = "none",plot.title = element_text(hjust = 0.5, face = "bold"), axis.title = element_text(size = 10)) 
```

As we can see, the less the price is, the more reviews there is. Cheap AirBnB are more attractive.

## Step 4 : Modeling and Prediction (Data Science)

### Linear Regression Model

-   A linear regression model was trained to predict the price (log-transformed) as a function of neighborhood (neighbourhood_group), room type (room_type) and number of reviews (number_of_reviews).

```{r}
# Modèle linéaire
lm_model = lm(log1p(price) ~ neighbourhood_group + room_type + number_of_reviews, data = Clean_train)

summary(lm_model)
```

### Decision Tree Model

-   A regression tree was built using rpart to visualize how decisions on price are structured.

```{r, include = FALSE}
library(rpart)
library(rpart.plot)
```

```{r}
tree_model <- rpart(price_log ~ neighbourhood_group + room_type + number_of_reviews, data = Clean_train, method = "anova")

rpart.plot(tree_model)
```

Now we see that 31% of people take shared or private room that cost 66\$ and 17% for 89\$ And 25% of people take home or apartment for 147\$ and 27% of them take it for 200\$

### Model Evaluation

-   Predictions were transformed back to the original scale and RMSE was calculated to evaluate model accuracy.

```{r}
# Prédictions
lm_pred <- predict(lm_model, Clean_train)
tree_pred <- predict(tree_model, Clean_train)

# Retour à l’échelle originale
lm_pred_exp <- expm1(lm_pred)
tree_pred_exp <- expm1(tree_pred)

# Calcul RMSE
lm_rmse <- sqrt(mean((Clean_train$price - lm_pred_exp)^2))
tree_rmse <- sqrt(mean((Clean_train$price - tree_pred_exp)^2))

lm_rmse
tree_rmse
```

Both models have an average error of about \$230, which is reasonable given the high variability of Airbnb prices in New York.

### Interpretation

The model shows that Airbnb prices in New York City vary significantly depending on the neighborhood and type of accommodation. Locations in Manhattan and Brooklyn are, on average, much more expensive than those in the Bronx. Conversely, "Private Rooms" and "Shared Rooms" are considerably cheaper than "Entire Homes." The number of reviews has a slight negative effect on price, suggesting that older or more popular listings tend to be slightly less expensive. Overall, these variables explain approximately 45% of the log-transformed price variation, making this a simple yet relevant model for an initial analysis of pricing factors.