---
title: "Project R"
author: "Legrand, Hu, Stintzi"
date: "2025-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project NYC Airbnb - Analyzing the Rental Market

Your data analyst team has been hired to study the Airbnb market in New York City. Using a public dataset, you must analyze price trends, the geographic distribution of listings, and the factors that influence rental prices. The project will culminate in the creation of a regression model to predict the price of a rental and an interactive map to visualize your findings.

```{r, include=FALSE}
train = read.csv("train.csv")
library(lubridate)
library(dplyr)
```

Handling Temporal Data: The last_review column is in text format. Convert it to a proper date format using the lubridate package.

```{r}
head(train)
```

```{r}
train$last_review = ymd(train$last_review)
train
```

 Treating Informative NA s: The last_review and reviews_per_month columns have NA s for listings with no reviews. Replace the .NA s in 
reviews_per_month with a proper number.

```{r}
train$reviews_per_month[is.na(train$reviews_per_month)] = 0
train
```

```{r}
CleanTrainPrice = train %>%
  filter(price != 0)
CleanTrainPrice
```

```{r}
Clean_train = CleanTrainPrice %>%
  select(-c(id, host_id, host_name))
Clean_train
```

## Step 2:

```{r}
Clean_train = Clean_train %>%
  mutate(price_log = log1p(price))
```


Verify the difference between price and price_log

```{r}
summary(Clean_train$price)
summary(Clean_train$price_log)
```

```{r}
# Calcul de la date la plus récente dans le dataset
max_date = max(Clean_train$last_review, na.rm = TRUE)

# Création de la nouvelle colonne
Clean_train = Clean_train %>%
  mutate(days_since_last_review = as.numeric(difftime(max_date, last_review, units = "days")))

# Vérification
Clean_train
```

```{r}
# Aperçu de la répartition des quartiers
table(Clean_train$neighbourhood_group)

# Et du nombre de quartiers détaillés
length(unique(Clean_train$neighbourhood))
```

```{r}
Clean_train$room_type = as.factor(Clean_train$room_type)

str(Clean_train$room_type)
levels(Clean_train$room_type)
```

